---
title: "MATH4753Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{MATH4753 Package Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
library(MATH4753)
```

Introduction

The MATH4753 package provides several statistical and probability functions used in OU's MATH 4753 course.
This vignette demonstrates four functions from the package—birthday(), mysum(), scatterhist(), and mybin()—and shows how to work with the included ddt dataset.

1. The birthday() Function

The birthday() function computes the probability that at least two people in a group share a birthday.

```{r}
birthday(20)
birthday(1:30)
```

2. The mysum() Function

mysum() returns the sum of integers from 1 to n.

```{r}
mysum(10)
mysum(25)
```

3. The scatterhist() Function

scatterhist() creates a scatterplot with marginal histograms.
Below is an example using randomly generated data.


```{r}
set.seed(123)
x <- rnorm(100)
y <- rnorm(100)
scatterhist(x, y)
```

4. The mybin() Function

mybin() computes the binomial probability for given n, p, and k.


```{r}
mybin(iter = 1000, n = 10, p = 0.5)
mybin(iter = 1000, n = 20, p = 0.25)
```

5. Dataset Example: ddt

The package includes the ddt dataset, which contains environmental measurement data.
We load and explore the dataset below.

```{r}
data(ddt)
head(ddt)
summary(ddt)
```

can see a simple relationship between variables:


```{r}
data(ddt)

head(ddt)
summary(ddt)

plot(ddt$DDT,
     main = "DDT Concentration Values",
     xlab = "Observation Index",
     ylab = "DDT Concentration")
```

Conclusion

This vignette demonstrated four major functions from the package and showed how to load and explore the included ddt dataset.
The package provides useful statistical tools for probability, simulation, and visualization.


---
